First steps in synthetic algebraic geometry

### 2017-03-20 snippets/injective.tex
typo

Since flabbiness is a local property, it's not unreasonable to expect that
flabbiness could be characterized in the internal language. The following
proposition shows that this is indeed the case.

\begin{prop}\label{prop:internal-char-flabbiness}
Let~$\F$ be a sheaf of sets on a topological space (or locale~$X$).
Then the following statements are equivalent:
\begin{enumerate}
\item $\F$ is flabby.

\item For any open subset $U \subseteq X$ and any section $s \in \F(U)$ there is
an open covering $X = \bigcup_i V_i$ such that, for each $i$, there is an
extension of $s$ to $U \cup V_i$ (that is, a section $s' \in \F(U \cup V_i)$
such that $s'|_U = s$).

@@

\begin{proof}
The implication~``1~$\Rightarrow$~2'' is trivial. The converse direction uses a
typical argument with Zorn's lemma, considering a maximal extension. The
equivalence~``$2~\Leftrightarrow~3$'' is routine, using the Kripke--Joyal
semantics to interpret the internal statement. Condition~4 is a straightforward
reformulation of Condition~3.
\end{proof}

@@

Recall that an object~$I$ of a category~$\C$ is \emph{injective} if and only if,
for any monomorphism~$X \to Y$ in~$\C$ and any morphism~$I \to X$, there
is a lifting such that the diagram
\[ \xymatrix{
  X \ar@{^{(}->}[r]\ar[d] & Y \ar@{-->}[dl] \\
  I
} \]
commutes. Equivalently, an object~$I$ is injective if and only if the Hom
functor~$\Hom_\C(\placeholder, I) : \C^\op \to \Set$ maps monomorphisms in~$\C$ to
surjective maps. This general definition is often specialized to one these cases:
to the category of modules over a ring, to the category of set-valued sheaves
on a topological space, and to the category of sheaves of~$\O_X$-modules on a
ringed space~$(X,\O_X)$.

The definition is seldomly applied in the category of sets, since in a classical
context it's easy to show that a set is injective if and only if it's
inhabited, thereby completely settling the question of which objects are
injective in a trivial manner.

The question is more interesting in an intuitionistic setting, since
intuitionistically one cannot prove that inhabited sets are
injective~\cite{aczel-et-al:injective}; but one can still verify that any set embeds
into an injective set. (The powerset~$\P(X)$ and even the smaller
set~$\P_{\leq1}(X)$ of subsingletons of a given set~$X$ are injective. We verify
this in the proof of Lemma~\ref{lemma:enough-flabby}.)

@@

\begin{prop}\label{prop:notions-of-internal-injectivity}
Let~$\E$ be an elementary topos with a natural numbers object. Then
the following statements about an object~$I \in \E$ are equivalent.
\begin{enumerate}
\item $I$ is internally injective.
\item The functor~$[\placeholder, I] : \E^\op \to \E$ maps monomorphisms in $\E$
to morphisms for which any global element of the target locally (after change of
base along an epimorphism) possesses a preimage.
\item For any morphism $p : A \to 1$ in $\E$, the object $p^*I$ has property~(1)
as an object of $\E/A$.
\item For any morphism $p : A \to 1$ in $\E$, the object $p^*I$ has property~(2)
as an object of $\E/A$.
\item From the point of view of the internal language of~$\E$, the object~$I$
is injective.\footnote{We have only introduced the internal language for sheaf
toposes. The general definition is in~\cite[Section~XXX]{shulman:stack}.}
\end{enumerate}
\end{prop}

@@

\begin{proof}By Theorem~\ref{thm:char-injectivity} respectively
Theorem~\ref{thm:char-injectivity-modules}, Injectivity can be characterized in the
internal language. Any such property is local.\end{proof}

@@

Thus let a subsingleton~$K \subseteq [F,I]$ be given. We consider the
injectivity diagram
\[ \xymatrix{
  F' \ar[r]\ar[d] & F \ar@{-->}[ld] \\
  I,
} \]
where~$F' \defeq \{ s \? F \,|\, \text{$K$ is inhabited} \} \subseteq F$ and the
solid vertical map sends~$s \in F'$ to~$f(s)$, where~$f$ is an arbitrary element
of~$K$. This association is well-defined. Since~$I$ is injective, a dotted lift
as indicated exists. If~$K$ is inhabited, this lift is an element of~$K$.

The same kind of proof works for the case of sheaves of modules, relying on
Theorem~\ref{thm:char-injectivity-modules} and substituting the definition
of~$F'$ by~``$F' \defeq \{ s \? F \,|\, \text{$s = 0$ or $K$ is inhabited}
\}$''.
\end{proof}

@@

\begin{lemma}\label{lemma:enough-flabby}
Let~$X$ be a topological space (or a locale). Any sheaf of sets
over~$X$ can be embedded into a injective (therefore flabby) sheaf of sets.
\end{lemma}

@@

\begin{rem}The \emph{Godement construction} is a well-known way of embedding an
inhabited sheaf of sets~$\F$ into an injective sheaf, namely embedding it into
the sheaf of not necessarily continuous sections of the étale space of~$\F$:
\[ U \subseteq X \quad\longmapsto\quad
  \prod_{x \in U} \F_x. \]
The sheaf~$\P_{\leq1}(\F)$ does not coincide with this construction.
Instead, it is the sheaf with
\[ U \subseteq X \quad\longmapsto\quad
  \{ \langle V, s \rangle \,|\,
    \text{$V \subseteq U$ open, $s \in \F(V)$} \}. \]
It's not possible to describe the Godement construction in the internal language
of~$\Sh(X)$, since in the Godement construction the underlying set of~$X$
enters. But the sheaf topos of~$X$ doesn't remember this set. For instance, if~$X$
is an inhabited indiscrete topological spac, then~$\Sh(X)$ is equivalent
to~$\Set$.
\end{rem}

>>>>

Since flabbiness is a local property, it's not unreasonable to expect that
flabbiness can be characterized in the internal language. The following
proposition shows that this is indeed the case.

\begin{prop}\label{prop:internal-char-flabbiness}
Let~$\F$ be a sheaf of sets on a topological space~$X$ (or a locale).
Then the following statements are equivalent:
\begin{enumerate}
\item $\F$ is flabby.

\item ``Any section of~$\F$ can be locally extended'':
For any open subset $U \subseteq X$ and any section $s \in \F(U)$ there is
an open covering $X = \bigcup_i V_i$ such that, for each $i$, there is an
extension of $s$ to $U \cup V_i$ (that is, a section $s' \in \F(U \cup V_i)$
such that $s'|_U = s$).

@@

\begin{proof}
The implication~``(1)~$\Rightarrow$~(2)'' is trivial. The converse direction uses a
typical argument with Zorn's lemma, considering a maximal extension. The
equivalence~``(2)~$\Leftrightarrow$~(3)'' is routine, using the Kripke--Joyal
semantics to interpret the internal statement. Condition~4 is a straightforward
reformulation of Condition~3.
\end{proof}

@@

Recall that an object~$I$ of a category~$\C$ is \emph{injective} if and only if,
for any monomorphism~$X \to Y$ in~$\C$ and any morphism~$X \to I$, there
is a lifting such that the diagram
\[ \xymatrix{
  X \ar@{^{(}->}[r]\ar[d] & Y \ar@{-->}[dl] \\
  I
} \]
commutes. Equivalently, an object~$I$ is injective if and only if the Hom
functor~$\Hom_\C(\placeholder, I) : \C^\op \to \Set$ maps monomorphisms in~$\C$ to
surjective maps. This general definition is often specialized to one of these cases:
to the category of modules over a ring, to the category of set-valued sheaves
on a topological space, and to the category of sheaves of~$\O_X$-modules on a
ringed space~$(X,\O_X)$.

The definition is seldomly applied in the category of sets, since in a classical
context it's easy to show that a set is injective if and only if it's
inhabited, thereby completely settling the question which objects are
injective in a trivial manner.

The question is more interesting in an intuitionistic setting, since
intuitionistically one cannot prove that inhabited sets are
injective~\cite{aczel-et-al:injective}; but one can still verify that any set embeds
into an injective set: The powerset~$\P(X)$ and even the smaller
set~$\P_{\leq1}(X)$ of subsingletons of a given set~$X$ are injective. We verify
this in the proof of Lemma~\ref{lemma:enough-flabby}.

@@

\begin{prop}\label{prop:notions-of-internal-injectivity}
Let~$\E$ be an elementary topos with a natural numbers object. Then
the following statements about an object~$I \in \E$ are equivalent.
\begin{enumerate}
\item $I$ is internally injective.
\item The functor~$[\placeholder, I] : \E^\op \to \E$ maps monomorphisms in $\E$
to morphisms for which any global element of the target locally (after change of
base along an epimorphism) possesses a preimage.
\item For any morphism $p : A \to 1$ in $\E$, the object $p^*I$ has property~(1)
as an object of $\E/A$.
\item For any morphism $p : A \to 1$ in $\E$, the object $p^*I$ has property~(2)
as an object of $\E/A$.
\item From the point of view of the internal language of~$\E$, the object~$I$
is injective.\footnote{In Section~\ref{sect:internal-language}, we have only
introduced the internal language for sheaf toposes. The general definition is
in~\cite[Section~7]{shulman:stack}.}
\end{enumerate}
\end{prop}

@@

\begin{proof}By Theorem~\ref{thm:char-injectivity} respectively
Theorem~\ref{thm:char-injectivity-modules}, injectivity can be characterized in the
internal language. Any such property is local.\end{proof}

@@

Thus let a subsingleton~$K \subseteq [F,I]$ be given. We consider the
injectivity diagram
\[ \xymatrix{
  F' \ar[r]\ar[d] & F \ar@{-->}[ld] \\
  I
} \]
where~$F'$ is the subset~$\{ s \? F \,|\, \text{$K$ is inhabited} \} \subseteq F$ and the
solid vertical map sends~$s \in F'$ to~$f(s)$, where~$f$ is an arbitrary element
of~$K$. This association is well-defined. Since~$I$ is injective, a dotted lift
as indicated exists. If~$K$ is inhabited, this lift is an element of~$K$.

The same kind of argument applies to the case of sheaves of modules, relying on
Theorem~\ref{thm:char-injectivity-modules} and defining~$F'$ as the
submodule~``$\{ s \? F \,|\, \text{$s = 0$ or $K$ is inhabited} \}$''.
\end{proof}

@@

\begin{lemma}\label{lemma:enough-flabby}
Let~$X$ be a topological space (or a locale). Any sheaf of sets
over~$X$ can be embedded into an injective (therefore flabby) sheaf of sets.
\end{lemma}

@@

\begin{rem}The \emph{Godement construction} is a well-known way of embedding an
inhabited sheaf of sets~$\F$ into an injective sheaf, namely embedding it into
the sheaf of not necessarily continuous sections of the étale space of~$\F$:
\[ U \subseteq X \quad\longmapsto\quad
  \prod_{x \in U} \F_x. \]
The sheaf~$\P_{\leq1}(\F)$ does not coincide with this construction.
Instead by Definition~\ref{defn:interpretation-internal-constructions}, it is the sheaf with
\[ U \subseteq X \quad\longmapsto\quad
  \{ \langle V, s \rangle \,|\,
    \text{$V \subseteq U$ open, $s \in \F(V)$} \}. \]
It's not possible to describe the Godement construction in the internal language
of~$\Sh(X)$, since in the Godement construction the underlying set of~$X$
enters. But the sheaf topos of~$X$ doesn't remember this set. For instance, if~$X$
is an inhabited indiscrete topological spac, then~$\Sh(X)$ is equivalent
to~$\Set$.
\end{rem}
